---
title: "MutationAnalysis"
author: "Mona Behrouzian, 50896695"
date: "13/11/2023"
output: pdf_document
---

```{r setup, include=FALSE}
#Set up code
knitr::opts_chunk$set(echo = TRUE)

#code from TA to make sure my commented code does not fall off the PDF page
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)

setwd("C:\\Users\\monab\\Documents\\R\\TCGA-KIRC")

#use this code to set the correct column names in importing
clinical_data <- read.table("data_clinical_patient.txt", fill = TRUE, stringsAsFactors = FALSE, header = TRUE, sep = '\t')
mutation_data <- read.table("data_mutations.txt", fill = TRUE, stringsAsFactors = FALSE, header = TRUE, sep = '\t')
RNA_data <- read.csv("RNAseq_KIRC.csv")
```

Trimming/cleaning the data
```{r}
#install.packages("dplyr")
library(dplyr)
library(stringr)

#clinical data set
length(unique(clinical_data$PATIENT_ID))


#mutation data set
mutation_data <- mutation_data %>% 
  mutate(Tumor_Sample_Barcode = str_remove(Tumor_Sample_Barcode, "-01"))

length(unique(mutation_data$Tumor_Sample_Barcode))


#Change the dots to dashes for the col names in RNAseq
#note we are useing "regular expression" syntax, where \\. is a dot and # is a dash
names(RNA_data) <- gsub(x = names(RNA_data), pattern = "\\.", replacement = "\\-")  

# deleting the excess characters in the column names
# ie keeping only up until the 12th character
names(RNA_data) <- substr(x = names(RNA_data), 1, 12)

length(unique(colnames(RNA_data)))

# RNA_seq_patients <- RNA_seq_patients %>% 
#   mutate(RNA_seq_patients = str_remove(RNA_seq_patients, "-01"))
```


```{r}
# Install pheatmap 
# install.packages("pheatmap")
library(ggplot2)
library(pheatmap)
```

### Plotting MAF summary
The summary of the maf file can be visualized through a stacked barplot depicting the number of variants in each sample, alongside a boxplot summarizing variant types based on Variant_Classification, and so on.

```{r}
#hugo symbol is gene names, "table" lets us summarize the content
hugo <- as.data.frame(table(mutation_data$Hugo_Symbol))
#output is one column of the gene name, and second column is frequency of each gene
#so now we know we have only 6,732 unique genes 

#again, variant_classification is the outcome or consequence of the mutation 
#output is the frequency of each mutation type
var.class <- as.data.frame(table(mutation_data$Variant_Classification))
#thus we have 17 types of variant classification 
#for eg, row 10 tells us how often we have a Missense Mutation (5259), meaning that the mutated codon changes the resulting amino acid

#plotting the variant classifications 
ggplot(data=var.class, aes(x=Var1, y=Freq))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
#the third line lets us plot the x axis names at an angle
```
# Variant Class
```{r}
var.class2 <- as.data.frame(table(mutation_data$VARIANT_CLASS))

ggplot(data=var.class2, aes(x=Var1, y=Freq))+
  geom_col(aes(fill=Var1))
#fill=var1 to have a different colour for each bar 
```

# Variant Type 
```{r}
var.type <- as.data.frame(table(mutation_data$Variant_Type))
ggplot(data=var.type, aes(x=Var1, y=Freq))+
  geom_col( aes(fill=Var1))
```

```{r}
sample.name <- as.data.frame(table(mutation_data$Tumor_Sample_Barcode))
#thus we ahve 354 patients! 

hugo <- as.data.frame(table(mutation_data$Hugo_Symbol))

#want to order it, so in the future, we can just analysize the top few genes
#code orders from highest to lowest frequency 
hugo.ordered <- hugo[order(-hugo$Freq),]

ggplot(data=hugo.ordered[1:15,], aes(x=Var1, y=Freq))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45,hjust=1))+
  scale_x_discrete(limits = hugo.ordered[1:15,]$Var1)
#the 4th line is important! otherwise r will sort alphabetically instead of highest to lowest. 
#we are saying to r: sort it based on the info i give you. not alphabetically 
```
graph shows top 15 mutations.
thus, the most mutated gene is VHL. ~100 mutations
but a majority of the mutated genes only occurred about 3 times. 

we are interested in the top genes with the highest mutation but we need a rational for how many top genes to select. so important to plot! 
visually, we can see the top 3 show significant more mutation. so 3 is a good start. this will likely give us more reliable clustering results.


### Generate oncoplot matrix

typically, oncoplots are based on VARIANT CLASSIFICATION

rows = genes
col = patients 

need to get unique variant classification 
```{r}
#install.packages("reshape2")
library(reshape2)
library(ggplot2)
library(pheatmap)

cnv_events = unique(mutation_data$Variant_Classification)

#dcast lets us reshape our data 
#because we have a longggg data set of rows 
oncomat = reshape2::dcast(
  data = mutation_data,
  
  #ie rows are hugo symbols, col are patients 
  formula = Hugo_Symbol ~ Tumor_Sample_Barcode, 
  
  #does a few complex things. converts everthing to char
  #some genes have more than 1 mutation. so this function tries to find ...??? 
  fun.aggregate = function(x, cnv = cnv_events) {
    x = as.character(x) # >= 2 same/distinct variant classification = Multi_Hit
    xad = x[x %in% cnv]
    xvc = x[!x %in% cnv]
    
    #if the number of mutation per that gene for that patient, say it is a multi-hit, otherwise keep it
    if (length(xvc) > 0) {
      xvc = ifelse(test = length(xvc) > 1,
                   yes = 'Multi_Hit',
                   no = xvc)
    }
    
    #??
    x = ifelse(
      test = length(xad) > 0,
      yes = paste(xad, xvc, sep = ';'),
      no = xvc
    )
    
    #these are the clean our real dataset
    x = gsub(pattern = ';$',
             replacement = '',
             x = x)
    x = gsub(pattern = '^;',
             replacement = '',
             x = x)
    return(x)
  },
  value.var = 'Variant_Classification',
  fill = '',
  drop = FALSE
)
```
output of this is oncomat variable
recall we had 17 types of mutation in this dataset, ie 17 variant classification types
so now we need to convert this into binary 

now its our choice.
if you want to make a binary matrix just based on missense mutation, then can just keep the missense as 1, convert everything to NA
or, could covert ALL type of mutation to 1

why do we need to do this? because input for clustering is numeric. must convert our char to binary

```{r}
rownames(oncomat) = oncomat$Hugo_Symbol
oncomat <- oncomat[,-1]

oncomat.ordered <- oncomat[order(-hugo$Freq),]
```
lots of data, ploting would be a big mess
so, its important for us to plot the highly mutated genes
earlier we decided on top 3
based on the frequency of hugo, we are ordering the oncomat 
now, you can see that the first row is VHL, which was the highest mutated gene we found earlier 

now going to convert it.
all the mutations are important for us. that is the default idea
BUT FOR THE PROJECT MUST FIND A HYPOTHESIS FOR WHICH VAR CLASS IS IMPORTANT
```{r}
mat <- oncomat.ordered
mat[mat!=""]=1
mat[mat==""]=0
mat <- apply(mat, 2 ,as.numeric)
mat <- as.matrix(mat)
rownames(mat)  <-  row.names(oncomat.ordered)
```
now can run clustering!

why do we need to run clustering on the oncoplot? 
because we want to group the patients based on the mutation types!!!
because we dont have a prior knowledge of mutation! 
sometimes we will know the hypothesis, eg itll be male vs female
but here, we dont know. trying to find any interesting infomation that can cluster the patients

### Draw the heatmap and cluster the patients
Patients can be grouped into clusters based on their most prevalent mutated gene or genes (here I selected the top 3 genes).
```{r}
library(pheatmap)

reduce.mat <- mat[1:3,]
res <- pheatmap(reduce.mat,
         cluster_rows = F,
         show_colnames=FALSE)
```
the rows are the top 3 mutated 
the columns are the patients 
so, based on the first layer we have 2 groups

get the cluster value
```{r}
cluster <-  as.data.frame(cutree(res$tree_col, k = 2))
#ie want to group people into 2 groups
cluster
```

now we want to do survival analysis. why? 
bcause grouping isnt enough. we need to clinically show that there is a difference between group 1 and group 2.
IF there is a clinical difference, then it probably was a good discovery. 

